---
title: "Worldbank population wrangling"
author: "Allison Horst"
date: "9/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
```

Get data:
```{r}
wb_pop <- read_csv(here("data","world_bank_pop.csv"))

# Explore the data and ask: how are the data NOT tidy? 
```

We have one clear issue that make these data *not* tidy. The year should be a single variable (column), but it is split up by individual years. We *may* consider total and urban populations separate variables (this one is a bit of a gray area). 

We will use pivot_longer() and pivot_wider() from the `tidyr` package to help us reshape this data into tidy format. 

First, by gathering the year data with pivot_longer():

```{r}
wb_pop_long <- wb_pop %>% 
  pivot_longer(cols = '2000':'2017', 
               names_to = "year",
               values_to = "population")
```

Then, we'll split the two population measures into separate columns using `pivot_wider()`, then clean up the names, then make sure year is numeric:

```{r}
wb_pop_tidy <- wb_pop_long %>% 
  pivot_wider(names_from = indicator, 
              values_from = population) %>% 
  clean_names() %>% 
  mutate(year = as.numeric(year))
```

## Git Check! 

Stage, commit, PULL, then push to GitHub. Ensure that the updates are stored by refreshing your repo on GitHub! Then come back to RStudio.

### Wrangling & graphs 

Now we have our data in tidy format, and have made sure our updates are safely stored, let's do some wrangling & make some plots! 

```{r}
# Filter to China (CHN), India (IND), USA (USA), Indonesia (IDN), Brazil (BRA)

top5_pop <- wb_pop_tidy %>% 
  filter(country %in% c("CHN", "IND", "USA", "IDN", "BRA"))
```

Then make a graph for each country's total population over time: 
```{r}
ggplot(data = top5_pop, aes(x = year, y = total_population)) +
  geom_line(aes(color = country))
```

### Git Check

Knit (& save), stage, commit, PULL, then push changes to GitHub. 

Go to GitHub & refresh repo to ensure it's updated. 

Come back to RStudio, CLOSE project & restart R. Reopen the project, PULL to make sure it's updated, then you can keep working. 

### END lesson! 
